\textbf{Microbiota predict \emph{Clostridium difficile} severity in germ-free mice colonized with human feces}

### 2016-12-07 submission/Manuscript.tex

\emph{Clostridium difficile} is an opportunistic pathogen of the human
lower gastrointestinal tract. \emph{C. difficile} forms spores that can
persist on abiotic surfaces and are not readily killed by ethanol-based
hand-sanitizers, putting hospital patients particularly at risk. Indeed,
\textasciitilde{}12\% of hospital acquired infections in the United
States are due to \emph{C. difficile} and result in up to 15,000 deaths
annually {[}1{]}. Disruption of the native microbial community is the
most common risk factor for development of \emph{C. difficile} infection
(CDI) {[}2{]}. Antibiotic use and inflammatory bowel diseases are
associated with loss of colonization resistance to \emph{C. difficile}
through the loss of potentially protective bacterial families such as
\emph{Barnesiella} and \emph{Lachnospiraceae} in both mouse models and
human association experiments {[}3,4{]}. The composition of the
community is clearly important for the acquisition, resistance and
treatment of \emph{C. difficile} infection, as giving patients a healthy
fecal microbiome transplant is the most effective treatment for this
disease (cite Anna 2014). The precise mechanisms of colonization
resistance and restoration of a healthy community are yet to be
discovered.

Murine models to study CDI typically rely on treating
conventionally-raised mice with antibiotics either in drinking water or
by injection to induce susceptibility (3, 4). This model provides a
convenient way to study \emph{C. difficile} pathogenesis and virulence
factors. Numerous microbiome studies have been performed using this
model to determine the antibiotic classes (5), starting microbial
community (6) and metabolites (7) that impact development and severity
of \emph{C. difficile} infection. While informative, these studies are
somewhat removed from human disease because they only examine
mouse-associated microbial communities.

Gnotobiotic or germ-free mouse models have been used for a range of
studies of CDI, including assessment of species-specific interactions
between \emph{C. difficile} and competing microbial community members
(8), analysis of nutrient restriction (9), \emph{in vivo}
transcriptomics of \emph{C. difficile} and examination of host immune
response to CDI (10). Further, CDI therapeutics such as antibiotics and
fecal microbiota transplants have been tested extensively in a
gnotobiotic-piglet or piglet-to-gnotobiotic-mouse model of disease (11),
(12). Pigs have a longer digestive tract with components more similar to
humans than mice and are typically infected by strains typical in human
infection (13). However, the murine and porcine microbiomes typically do
not resemble those of the human gut.

The power of the gnotobiotic models to study CDI has been further
realized by first inoculating germ-free mice and piglets with human
stool microbes. In one study, germ-free piglets were acutely colonized
with human feces for one week and then treated with tigecycline. After
challenge with \emph{C. difficile} none of the antibiotic-treated
piglets succumbed to infection, while some of the untreated
human-colonized pigs did (11). In another study, germ-free mice
colonized with human feces were bred over several generations to create
a cohort of mice with identical human-derived microbiomes (14). These
mice were subsequently treated with a five-antibiotic cocktail to induce
dysbiosis and then were successfully colonized by \emph{C. difficile}
(14). While informative, these studies were limited in their use of only
one human donor as input inoculum.

@@

\textbf{Nick will be updating the next two sections} \textbf{Microbes
present in the gut prior to infection are predictive of \emph{C.
difficile} colonization levels} Previous work in our group has
demonstrated that the microbes present in the mouse microbiome prior to
\emph{C. difficile} challenge can predict future colonization levels
(cite Alyx). To determine if the human-associated microbiome in the
mouse could similarly predict \emph{C. difficile} colonization we
employed a Random Forest machine learning algorithm similar to the one
used in previous studies. Relative abundance of OTUs with greater than
1\% abundance at day 0 were used as input to the model. We found that
the model explained X\% of the variance in colonization levels of
\emph{C. difficile} (Fig 3A). By refining the model to include only the
top X predictive OTUs, the model explained X\% of the variance. Partial
dependency plots of the top X OTUs predictive of \emph{C. difficile}
colonization reveal that colonization is positively associated with
\emph{Bacteriodetes}, \emph{Parabacteriodes} and
\emph{Porphyromonadaceae} (really?) (Fig 3B). Colonization was
negatively associated with the presence of \emph{Lachnospiraceae} (Fig
3B) though none of the human-associated mouse communities were
completely resistant to colonization by \emph{C. difficile} (Fig 2A,
3B). These results confirm that as in mouse microbiome studies, the
human-associated gut community prior to challenge is predictive of
\emph{C. difficile} colonization levels.

\textbf{The microbiome predicts \emph{C. difficile} infection severity.}
As \emph{C. difficile} colonization levels can be predicted by the
microbiome composition on day 0, we postulated that severity of disease
could also be predicted from day 0. Using weight loss as a proxy for
severity, we built a Random Forest classification model based on
relative abundance to determine which community members were associated
with either mild or severe disease. We optimized the model using the
AUC-RF algorithm to generate a Reciever Operating Curve (ROC) with
maximal Area Under the Curve (AUC) (cite Calle 2011), as used previously
in our group to classify colon cancer outcomes (cite Niel 2015 gen med).
The optimized model used X number of OTUs and identified several that
predicted increased weightloss and subsequent disease severity (Fig 4A,
4B). The populations that were most commonly found in mice that
succumbed to \emph{C. difficile} infection were members of the
\emph{Clostridiales}, \emph{Bacilliales}, \emph{Streptococcus}
(Lactobacilles?) and \emph{Bacteriodes}. (need more OTUs here, or number
that were used for model and then group by rank). The AUC was calculated
using the Out-Of-Bag (OOB) error for each sample (Fig 4A).
Cross-validation using 10-fold, leave-one-mouse-out and
leave-one-cage-out resulted in AUCs not significantly different from the
OOB (Fig 4A). This model reveals that the levels of X OTUs on day 0 can
predict disease outcome in mice with human-associated microbiomes.
(probably also going to want a supplemental table for predictive OTUs
for this and Fig 3)

\textbf{Susceptibility to infection is community-dependent.} \emph{C.
difficile} strains are known to vary in disease phenotypes based both on
pathogenicity factors of the strain and host microbiome composition
(cite someone). To test the susceptibility of human-derived microbiomes
to \emph{C. difficile} infection, two donor-mouse microbiomes that were
previously shown to be resistant to \emph{C. difficile} strain 430
induced weight loss (Fig 2B) (DA00396, DA00430-wait isnt this a control)
and one that was susceptible to severe disease (DA00578) were
independently challenged with one of three unique \emph{C. difficile}
strains (\emph{table or reference for toxin prod}, etc). As done
previously, \emph{C. difficile} stool CFU was enumerated daily and mouse
weight-loss tracked. Infection of resistant mice with virulent strain
458 or the weaker strain 299 resulted in colonization to levels near
that of 431 (Fig 5A). Accordingly, these mice did not experience severe
weight loss or show clinical signs of disease (Fig 5B). In contrast,
infection of susceptible mice (DA00578, Fig 1B) with pathogenic strain
395 resulted in colonization and severe weight loss by two days
post-infection (Fig 5A, 5B). Strain 395 is similarly virulent as strain
431, and shows a similar infection phenotype in these mice. These
preliminary results suggest that human-associated microbiomes may be
more predictive of disease severity than the \emph{C. difficile} strain.

@@

Study of human-associated microbiomes and their interactions with
\emph{C. difficile} has been limited by the lack of an appropriate model
and available samples. Here we colonized 16 groups of mice with human
donor stool from patients with different clinical profiles or healthy
volunteers. This created a diverse set of human-associated mouse
microbiomes that allowed for testing \emph{C. difficile} infection
dynamics and community interactions. Our results showed that both
healthy and diarrheal-associated human communities were susceptible to
severe disease. Additionally, our results confirmed previous findings
that resistance to \emph{C. difficile} disease is dependent on the
entire microbial community and not just one or two taxa (cite alyx).
Further, we were able to learn classification models that identified
specific OTUs within susceptible communities that predispose mice to
infection. By subsequently infecting susceptible or resistant
communities with varying strains of \emph{C. difficile} we confirmed
that the initial state of the human-associated microbiome is more
predictive of disease severity than the strain used.

Here we colonized germ-free mice with a variety of human stool
communities and allowed these communities to colonize the mice for 2
weeks, prior to \emph{C. difficile} challenge. It is important to note
that the mouse gut environment is different from that of the human, and
as such exerts a selective pressure on the human stool microbial
community that stabilizes over the 2 weeks. Thus these mice cannot be
considered ``humanized'' but rather colonized by human-associated
microbes. Interestingly, clinical status was not associated with
proclivity to severe \emph{C. difficile} infection as some healthy donor
mice were susceptible. A surprising finding was that none of the
human-associated mouse communities were completely resistant to
colonization by \emph{C. difficile}. This is in contrast to numerous
reports of mouse microbiome studies (cite) and the knowledge that in
general healthy humans without gut perturbations are resistant to
\emph{C. difficile} infection (cite). On the other hand, it is known
that humans, especially health care employees, can be asymptomatic
carriers of \emph{C. difficile} (Furyura-Kanamari). This may explain the
fact that the majority of our groups of mice were colonized by \emph{C.
difficile} but did not display outward signs of illness or
infection-associated weight loss (Fig 2).

@@

Finally, our pilot experiments testing the efficiency of infection with
different \emph{C. difficile} strains have allowed for testing community
versus pathogen factors of infection. Our results are consistent with
previous reports that the composition of the gut microbiome determines
if an individual is successfully infected or not (cite Alyx and others).
\emph{C. difficile} strains are known to vary in pathogenicity,
primarily via the amount of toxin that is produced during infection
(cite). To confirm that disease risk is dependent on the microbiome, we
challenged susceptibile mice with an additional pathogenic strain of
\emph{C. difficile} and found that the mice indeed fell ill. In
contrast, we exposed an asymptomatic community to a relatively weak
strain of \emph{C. difficile} and found that the community could be
colonized by this one as well-- but not cause disease. This suggests
that this particular colony is prone to colonization by any \emph{C.
difficile} strain. Further experiments will test if infecting
susceptible communities with weak strains allows the mice to survive. It
has been recently reported that carriage of a non-toxigenic strain of
\emph{C. difficile} can prevent additional colonization by a pathogenic
strain, suggesting a role of niche-filling or immunity this
pre-colonization provides (cite cite). One interesting experiment would
be to recapitulate this finding in human-associated mouse microbiomes by
making a susceptible community resistant to infection. Further work with
human-associated communities in mice and mathematical modeling will
allow for uncovering of complex human-associated disease processes and
dynamics as they relate to human \emph{C. difficile} infection.

@@

\textbf{Human samples} Donor stool samples were chosen from a sample
bank collected from patients from October 2010 to November 2012 at the
University of Michigan. Sample collection methods and study was approved
by the Institutional Review Board at the University of Michigan. Stool
samples from University of Michigan Health System patients were
submitted to the microbiology lab for \emph{C. difficile} testing.
\emph{C. difficile} presence was measured using WHATEVER- Cdiff quik
check? toxin AB?. Samples from these patients were stored in Cary-Blair
medium at -80C. All patients consented to participation in the study.
Donor stool for this study was selected from one patient that tested
positive for \emph{C. difficile} infection and five patients that
presented with diarrhea and tested negative for \emph{C. difficile}.
Healthy donor stool was collected from adult volunteers living in the
Ann Arbor, MI area. 100ul of stool slurry was gavaged into each mouse.

\textbf{\emph{Clostridium difficile} and sample processing} For all
experiments, 100 spores of \emph{Clostridium difficile} were used to
orally infect mice as described previously (cite alyx). The \emph{C.
difficile} strains used were acquired from spore stocks of (WHERE) and
comprised of strains 431, 395, 299 and 468 (cite?). After gavage, the
remaining inoculum was diluted and plated to confirm infection dose.
Bacteria were cultured anaerobically on TCCFA plates and incubated at
37C (cite plate recipe). During infection, two fresh stool pellets were
collected from each mouse daily. One pellet was resuspended in anaerobic
phosphate-buffered saline and diluted and plated on TCCFA plates to
enumerate \emph{C. difficile} CFU. The other pellet was frozen at -80C
until the end of the experiment.

\textbf{DNA isolation and sequencing} Bacterial DNA was isolated from
banked donor stool and each mouse stool pellet using the MO-BIO
PowerSoil DNA isolation kit. This DNA was used as template for
amplification of the V4 region of the 16S rRNA gene and fragments were
sequenced on an Illumina MiSeq as previously described (cite Jim paper).

@@

We would like to thank members of the Schloss lab and Jhansi Leslie for
analysis advice. We would also like to thank the Microbial Systems
Laboratory at the University of Michigan for sequencing assistance.

@@

\textbf{Figure 5. Infection of mice with different \emph{C. difficile
strains}.} 3 strains of \emph{C. difficile} were used to infect mice
colonized with susceptible (DA00578) or resistant (DA00369, DA00430)
human donor stool. A) \emph{C. difficile} stool CFU was enumerated over
10 days. B) Percent weight loss was calculated each day for each mouse.
In both plots, each mouse is a point and lines represent the mean of
each cage.

@@

>>>>

\emph{Clostridium difficile} is an opportunistic pathogen of the human
lower gastrointestinal tract. \emph{C. difficile} forms spores that can
persist on abiotic surfaces and are not readily killed by ethanol-based
hand-sanitizers, putting hospital patients particularly at risk. Indeed,
\textasciitilde{}12\% of hospital acquired infections in the United
States are due to \emph{C. difficile} and result in up to 15,000 deaths
annually {[}1{]}. Disruption of the native microbial community is the
most common risk factor for development of \emph{C. difficile} infection
(CDI) {[}2{]}. Antibiotic use and inflammatory bowel diseases are
associated with loss of colonization resistance to \emph{C. difficile}
through the loss of potentially protective bacterial families such as
\emph{Barnesiella} and \emph{Lachnospiraceae} in both mouse models and
human association experiments {[}3,4{]}. The composition of the
community is clearly important for the acquisition, resistance and
treatment of \emph{C. difficile} infection, as giving patients a healthy
fecal microbiome transplant is the most effective treatment for this
disease {[}5{]}. The precise mechanisms of colonization resistance and
restoration of a healthy community are yet to be discovered.

Murine models to study CDI typically rely on treating
conventionally-raised mice with antibiotics either in drinking water or
by injection to induce susceptibility {[}6,7{]}. This model provides a
convenient way to study \emph{C. difficile} pathogenesis and virulence
factors. Numerous microbiome studies have been performed using this
model to determine the antibiotic classes {[}8{]}, starting microbial
community {[}9{]} and metabolites {[}10{]} that impact development and
severity of \emph{C. difficile} infection. While informative, these
studies are somewhat removed from human disease because they only
examine mouse-associated microbial communities.

Gnotobiotic or germ-free mouse models have been used for a range of
studies of CDI, including assessment of species-specific interactions
between \emph{C. difficile} and competing microbial community members
{[}11{]}, analysis of nutrient restriction {[}12{]}, \emph{in vivo}
transcriptomics of \emph{C. difficile} and examination of host immune
response to CDI {[}\textbf{???}{]}. Further, CDI therapeutics such as
antibiotics and fecal microbiota transplants have been tested
extensively in a gnotobiotic-piglet or piglet-to-gnotobiotic-mouse model
of disease {[}13,14{]}. Pigs have a longer digestive tract with
components more similar to humans than mice and are typically infected
by strains typical in human infection {[}15{]}. However, the murine and
porcine microbiomes typically do not resemble those of the human gut.

The power of the gnotobiotic models to study CDI has been further
realized by first inoculating germ-free mice and piglets with human
stool microbes. In one study, germ-free piglets were acutely colonized
with human feces for one week and then treated with tigecycline. After
challenge with \emph{C. difficile} none of the antibiotic-treated
piglets succumbed to infection, while some of the untreated
human-colonized pigs did {[}13{]}. In another study, germ-free mice
colonized with human feces were bred over several generations to create
a cohort of mice with identical human-derived microbiomes {[}16{]}.
These mice were subsequently treated with a five-antibiotic cocktail to
induce dysbiosis and then were successfully colonized by \emph{C.
difficile} {[}16{]}. While informative, these studies were limited in
their use of only one human donor as input inoculum.

@@

\textbf{Nick will be updating the next two sections} \textbf{Microbes
present in the gut prior to infection are predictive of \emph{C.
difficile} colonization levels} Previous work in our group has
demonstrated that the microbes present in the mouse microbiome prior to
\emph{C. difficile} challenge can predict future colonization levels
{[}8{]}. To determine if the human-associated microbiome in the mouse
could similarly predict \emph{C. difficile} colonization we employed a
Random Forest machine learning algorithm similar to the one used in
previous studies. Relative abundance of OTUs with greater than 1\%
abundance at day 0 were used as input to the model. We found that the
model explained X\% of the variance in colonization levels of \emph{C.
difficile} (Fig 3A). By refining the model to include only the top X
predictive OTUs, the model explained X\% of the variance. Partial
dependency plots of the top X OTUs predictive of \emph{C. difficile}
colonization reveal that colonization is positively associated with
\emph{Bacteriodetes}, \emph{Parabacteriodes} and
\emph{Porphyromonadaceae} (really?) (Fig 3B). Colonization was
negatively associated with the presence of \emph{Lachnospiraceae} (Fig
3B) though none of the human-associated mouse communities were
completely resistant to colonization by \emph{C. difficile} (Fig 2A,
3B). These results confirm that as in mouse microbiome studies, the
human-associated gut community prior to challenge is predictive of
\emph{C. difficile} colonization levels.

\textbf{The microbiome predicts \emph{C. difficile} infection severity.}
As \emph{C. difficile} colonization levels can be predicted by the
microbiome composition on day 0, we postulated that severity of disease
could also be predicted from day 0. Using weight loss as a proxy for
severity, we built a Random Forest classification model based on
relative abundance to determine which community members were associated
with either mild or severe disease. We optimized the model using the
AUC-RF algorithm to generate a Reciever Operating Curve (ROC) with
maximal Area Under the Curve (AUC) {[}17{]}, as used previously in our
group to classify colon cancer outcomes {[}18{]}. The optimized model
used X number of OTUs and identified several that predicted increased
weightloss and subsequent disease severity (Fig 4A, 4B). The populations
that were most commonly found in mice that succumbed to \emph{C.
difficile} infection were members of the \emph{Clostridiales},
\emph{Bacilliales}, \emph{Streptococcus} (Lactobacilles?) and
\emph{Bacteriodes}. (need more OTUs here, or number that were used for
model and then group by rank). The AUC was calculated using the
Out-Of-Bag (OOB) error for each sample (Fig 4A). Cross-validation using
10-fold, leave-one-mouse-out and leave-one-cage-out resulted in AUCs not
significantly different from the OOB (Fig 4A). This model reveals that
the levels of X OTUs on day 0 can predict disease outcome in mice with
human-associated microbiomes. (probably also going to want a
supplemental table for predictive OTUs for this and Fig 3)

@@

Study of human-associated microbiomes and their interactions with
\emph{C. difficile} has been limited by the lack of an appropriate model
and available samples. Here we colonized 16 groups of mice with human
donor stool from patients with different clinical profiles or healthy
volunteers. This created a diverse set of human-associated mouse
microbiomes that allowed for testing \emph{C. difficile} infection
dynamics and community interactions. Our results showed that both
healthy and diarrheal-associated human communities were susceptible to
severe disease. Additionally, our results confirmed previous findings
that resistance to \emph{C. difficile} disease is dependent on the
entire microbial community and not just one or two taxa {[}8{]}.
Further, we were able to learn classification models that identified
specific OTUs within susceptible communities that predispose mice to
infection.

Here we colonized germ-free mice with a variety of human stool
communities and allowed these communities to colonize the mice for 2
weeks, prior to \emph{C. difficile} challenge. It is important to note
that the mouse gut environment is different from that of the human, and
as such exerts a selective pressure on the human stool microbial
community that stabilizes over the 2 weeks. Thus these mice cannot be
considered ``humanized'' but rather colonized by human-associated
microbes. Interestingly, clinical status was not associated with
proclivity to severe \emph{C. difficile} infection as some healthy donor
mice were susceptible. A surprising finding was that none of the
human-associated mouse communities were completely resistant to
colonization by \emph{C. difficile}. This is in contrast to numerous
reports of mouse microbiome studies {[}8{]} and the knowledge that in
general healthy humans without gut perturbations are resistant to
\emph{C. difficile} infection {[}Heinlen2010{]}. On the other hand, it
is known that humans, especially health care employees, can be
asymptomatic carriers of \emph{C. difficile} {[}19{]}. This may explain
the fact that the majority of our groups of mice were colonized by
\emph{C. difficile} but did not display outward signs of illness or
infection-associated weight loss (Fig 2).

@@

Further experiments will test if infecting susceptible communities with
weak strains allows the mice to survive. It has been recently reported
that carriage of a non-toxigenic strain of \emph{C. difficile} can
prevent additional colonization or recurrence by a pathogenic strain,
suggesting a role of niche-filling or immunity this pre-colonization
provides {[}20{]}. One interesting experiment would be to recapitulate
this finding in human-associated mouse microbiomes by making a
susceptible community resistant to infection. Further work with
human-associated communities in mice and mathematical modeling will
allow for uncovering of complex human-associated disease processes and
dynamics as they relate to human \emph{C. difficile} infection.

@@

\textbf{Human samples} Donor stool samples were chosen from a sample
bank collected from patients from October 2010 to November 2012 at the
University of Michigan. Sample collection methods and study was approved
by the Institutional Review Board at the University of Michigan. Stool
samples from University of Michigan Health System patients were
submitted to the microbiology lab for \emph{C. difficile} testing.
\emph{C. difficile} presence was measured using C.Diff Quik Chek
Complete Assay (Techlab). Samples from these patients were stored in
Cary-Blair medium at -80C. All patients consented to participation in
the study. Donor stool for this study was selected from one patient that
tested positive for \emph{C. difficile} infection and five patients that
presented with diarrhea and tested negative for \emph{C. difficile}.
Healthy donor stool was collected from adult volunteers living in the
Ann Arbor, MI area. 100ul of stool slurry was gavaged into each mouse.

\textbf{\emph{Clostridium difficile} and sample processing} For all
experiments, 100 spores of \emph{Clostridium difficile} were used to
orally infect mice as described previously {[}8{]}. \emph{C. difficile}
strain 431 is an 027 ribotype clinical isolate isolated from a patient
with severe CDI {[}21{]}. After gavage, the remaining inoculum was
diluted and plated to confirm infection dose. Bacteria were cultured
anaerobically on TCCFA plates and incubated at 37C {[}22{]}. During
infection, two fresh stool pellets were collected from each mouse daily.
One pellet was resuspended in anaerobic phosphate-buffered saline and
diluted and plated on TCCFA plates to enumerate \emph{C. difficile} CFU.
The other pellet was frozen at -80C until the end of the experiment.

\textbf{DNA isolation and sequencing} Bacterial DNA was isolated from
banked donor stool and each mouse stool pellet using the MO-BIO
PowerSoil DNA isolation kit. This DNA was used as template for
amplification of the V4 region of the 16S rRNA gene and fragments were
sequenced on an Illumina MiSeq as previously described {[}23{]}.

@@

We would like to thank members of the Schloss lab and Jhansi Leslie for
experimental and analytical advice. We would also like to thank the
Microbial Systems Laboratory at the University of Michigan for
sequencing assistance.

@@

@@

5. Seekatz AM, Aas J, Gessert CE, Rubin TA, Saman DM, Bakken JS, et al.
Recovery of the gut microbiome following fecal microbiota
transplantation. mBio. American Society for Microbiology; 2014;5:
e00893--14--e00893--14.
doi:\href{http://dx.doi.org/10.1128/mbio.00893-14}{10.1128/mbio.00893-14}

6. Buffie CG, Jarchum I, Equinda M, Lipuma L, Gobourne A, Viale A, et
al. Profound alterations of intestinal microbiota following a single
dose of clindamycin results in sustained susceptibility to clostridium
difficile-induced colitis. Infection and Immunity. American Society for
Microbiology; 2011;80: 62--73.
doi:\href{http://dx.doi.org/10.1128/iai.05496-11}{10.1128/iai.05496-11}

7. Theriot CM, Koumpouras CC, Carlson PE, Bergin II, Aronoff DM, Young
VB. Cefoperazone-treated mice as an experimental platform to assess
differential virulence of clostridium difficile strains. Gut Microbes.
Informa UK Limited; 2011;2: 326--334.
doi:\href{http://dx.doi.org/10.4161/gmic.19142}{10.4161/gmic.19142}

8. Schubert AM, Sinani H, Schloss PD. Antibiotic-induced alterations of
the murine gut microbiota and subsequent effects on colonization
resistance against clostridium difficile. mBio. American Society for
Microbiology; 2015;6: e00974--15.
doi:\href{http://dx.doi.org/10.1128/mbio.00974-15}{10.1128/mbio.00974-15}

9. Reeves AE, Theriot CM, Bergin IL, Huffnagle GB, Schloss PD, Young VB.
The interplay between microbiome dynamics and pathogen dynamics in a
murine model of clostridium difficile infection. Gut Microbes. Informa
UK Limited; 2011;2: 145--158.
doi:\href{http://dx.doi.org/10.4161/gmic.2.3.16333}{10.4161/gmic.2.3.16333}

10. Theriot CM, Young VB. Microbial and metabolic interactions between
the gastrointestinal tract and clostridium difficile infection. Gut
Microbes. Informa UK Limited; 2013;5: 86--95.
doi:\href{http://dx.doi.org/10.4161/gmic.27131}{10.4161/gmic.27131}

11. Reeves AE, Koenigsknecht MJ, Bergin IL, Young VB. Suppression of
clostridium difficile in the gastrointestinal tracts of germfree mice
inoculated with a murine isolate from the family lachnospiraceae.
Infection and Immunity. American Society for Microbiology; 2012;80:
3786--3794.
doi:\href{http://dx.doi.org/10.1128/iai.00647-12}{10.1128/iai.00647-12}

12. Ng KM, Ferreyra JA, Higginbottom SK, Lynch JB, Kashyap PC, Gopinath
S, et al. Microbiota-liberated host sugars facilitate post-antibiotic
expansion of enteric pathogens. Nature. Springer Nature; 2013;502:
96--99.
doi:\href{http://dx.doi.org/10.1038/nature12503}{10.1038/nature12503}

13. Kim HB, Zhang Q, Sun X, Beamer G, Wang Y, Tzipori S. Beneficial
effect of oral tigecycline treatment on clostridium difficile infection
in gnotobiotic piglets. Antimicrobial Agents and Chemotherapy. American
Society for Microbiology; 2014;58: 7560--7564.
doi:\href{http://dx.doi.org/10.1128/aac.03447-14}{10.1128/aac.03447-14}

14. Diao H, Yan HL, Xiao Y, Yu B, Yu J, He J, et al. Intestinal
microbiota could transfer host gut characteristics from pigs to mice.
BMC Microbiology. Springer Nature; 2016;16.
doi:\href{http://dx.doi.org/10.1186/s12866-016-0851-z}{10.1186/s12866-016-0851-z}

15. Steele J, Feng H, Parry N, Tzipori S. Piglet models of acute or
chronic clostridium difficile illness. The Journal of Infectious
Diseases. Oxford University Press (OUP); 2010;201: 428--434.
doi:\href{http://dx.doi.org/10.1086/649799}{10.1086/649799}

16. Collins J, Auchtung JM, Schaefer L, Eaton KA, Britton RA. Humanized
microbiota mice as a model of recurrent clostridium difficile disease.
Microbiome. Springer Nature; 2015;3.
doi:\href{http://dx.doi.org/10.1186/s40168-015-0097-2}{10.1186/s40168-015-0097-2}

17. Calle ML, Urrea V, Boulesteix A-L, Malats N. AUC-RF: A new strategy
for genomic profiling with random forest. Human Heredity. S. Karger AG;
2011;72: 121--132.
doi:\href{http://dx.doi.org/10.1159/000330778}{10.1159/000330778}

18. Baxter NT, Ruffin MT, Rogers MAM, Schloss PD. Microbiota-based model
improves the sensitivity of fecal immunochemical test for detecting
colonic lesions. Genome Medicine. Springer Nature; 2016;8.
doi:\href{http://dx.doi.org/10.1186/s13073-016-0290-3}{10.1186/s13073-016-0290-3}

19. Furuya-Kanamori L, Marquess J, Yakob L, Riley TV, Paterson DL,
Foster NF, et al. Asymptomatic clostridium difficile colonization:
Epidemiology and clinical implications. BMC Infectious Diseases.
Springer Nature; 2015;15.
doi:\href{http://dx.doi.org/10.1186/s12879-015-1258-4}{10.1186/s12879-015-1258-4}

20. Gerding DN, Meyer T, Lee C, Cohen SH, Murthy UK, Poirier A, et al.
Administration of spores of nontoxigenic clostridium difficile strain m3
for prevention of recurrent c difficile infection. JAMA. American
Medical Association (AMA); 2015;313: 1719.
doi:\href{http://dx.doi.org/10.1001/jama.2015.3725}{10.1001/jama.2015.3725}

21. Carlson PE, Walk ST, Bourgis AE, Liu MW, Kopliku F, Lo E, et al. The
relationship between phenotype, ribotype, and clinical disease in human
clostridium difficile isolates. Anaerobe. Elsevier BV; 2013;24:
109--116.
doi:\href{http://dx.doi.org/10.1016/j.anaerobe.2013.04.003}{10.1016/j.anaerobe.2013.04.003}

22. Bliss DZ, Johnson S, Clabots CR, Savik K, Gerding DN. Comparison of
cycloserine-cefoxitin-fructose agar (CCFA) and taurocholate-CCFA for
recovery of clostridium difficile during surveillance of hospitalized
patients. Diagnostic Microbiology and Infectious Disease. Elsevier BV;
1997;29: 1--4.
doi:\href{http://dx.doi.org/10.1016/s0732-8893(97)00113-2}{10.1016/s0732-8893(97)00113-2}

23. Kozich JJ, Westcott SL, Baxter NT, Highlander SK, Schloss PD.
Development of a dual-index sequencing strategy and curation pipeline
for analyzing amplicon sequence data on the MiSeq illumina sequencing
platform. Applied and Environmental Microbiology. American Society for
Microbiology; 2013;79: 5112--5120.
doi:\href{http://dx.doi.org/10.1128/aem.01043-13}{10.1128/aem.01043-13}
